\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-12-02 paper.qmd
Finalize the aspect rations

**tSNE**

```{r}
result_tsne <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_tsne_results.rds'))
```

```{r}
result_tsne |>
  kableExtra::kable(format = "latex",
                    booktabs = TRUE,
                    longtable = TRUE,
                    label = "scdeedtsne") |>
  kableExtra::kable_styling(latex_options = "scale_down")
```

```{r}
#| out-width: 50%

ggplot(data = result_tsne,
       aes(
         x = as.numeric(perplexity),
         y = number_dubious_cells
       )) +
  geom_line() +
  geom_point() +
  labs(x = "perplexity",
       y = "# of dubious cells") +
  theme_minimal()
```

@@

```{r}
#| out-width: 50%

@@

error_umap_plot_pbmc <- ggplot(error_pbmc, 
                               aes(x = a1, 
                                   y = MSE, 
                                   colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-height: 100%

free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2)
```

>>>>

@@

```{r}
#| out-width: 50%
#| eval: false

@@

error_umap_plot_pbmc <- ggplot(error_pbmc, 
                               aes(x = a1, 
                                   y = MSE, 
                                   colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-height: 100%
#| fig-cap: "Assessing which of the 2 tSNE layouts with different perplexity ($30$ (red), $320$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents tSNE layout (a, b). Layout b is universally suboptimal. Layout b is the best choice."

free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) +
  plot_layout(ncol = 1, heights = c(1, 1))
```

```{r}
#| label: published-pbmc-scdeed

training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds")
training_data_pbmc <- training_data_pbmc[, 1:50] |>
  mutate(ID = 1:NROW(training_data_pbmc))

## For umap
umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
umap_pbmc <- as_tibble(umap_pbmc)
names(umap_pbmc) <- c("UMAP1", "UMAP2")
umap_pbmc <- umap_pbmc |>
  mutate(ID = 1:NROW(umap_pbmc))

pbmc_scaled_obj_umap <- gen_scaled_data(
  data = umap_pbmc)
umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr

lim1 <- pbmc_scaled_obj_umap$lim1
lim2 <- pbmc_scaled_obj_umap$lim2
r2_umap <- diff(lim2)/diff(lim1)

```{r}
#| label: hexbin-pbmc-scdeed
## Compute hexbin parameters
num_bins_x_pbmc <- 12
lim1 <- pbmc_scaled_obj_umap$lim1
lim2 <- pbmc_scaled_obj_umap$lim2
r2_pbmc <- diff(lim2)/diff(lim1)

trimesh_removed_pbmc <- ggplot() + 
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  geom_point(
    data = umap_pbmc_scaled,
    aes(
      x = UMAP1,
      y = UMAP2
    ),
    alpha = 0.2,
    color = '#a65628'
  )  +
  interior_annotation("a1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

### 2024-12-02 paper.tex
Finalize the aspect rations

\textbf{tSNE}

\begin{longtable*}{lr}
\toprule
perplexity & number\_dubious\_cells\\
\midrule
5 & 87\\
10 & 143\\
15 & 146\\
20 & 101\\
25 & 93\\
\addlinespace
30 & 83\\
35 & 82\\
40 & 85\\
45 & 71\\
50 & 41\\
\addlinespace
55 & 48\\
60 & 46\\
65 & 35\\
70 & 36\\
75 & 49\\
\addlinespace
80 & 46\\
85 & 75\\
90 & 82\\
95 & 62\\
100 & 20\\
\addlinespace
105 & 19\\
110 & 19\\
115 & 21\\
120 & 18\\
125 & 17\\
\addlinespace
130 & 15\\
135 & 17\\
140 & 17\\
145 & 18\\
\bottomrule
\end{longtable*}

\begin{center}
\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-62-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-65-1.pdf}
\end{center}

@@

\caption{\label{fig-pbmc-mse-umap}}

>>>>

@@

@@

\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 tSNE layouts
with different perplexity (\(30\) (red), \(320\) (blue)) on the PBMC
data is the better representation using MSE for varying binwidth
(\(a_1\)). Colour used for the lines and points in the left plot and in
the scatterplots represents tSNE layout (a, b). Layout b is universally
suboptimal. Layout b is the best choice.}
